### 2022-09-23 chapters/optimization/python-dir/python_interperters.tex
corrections : end part
fix: missing references

      \item \textbf{Parakeet}~\cite{DBLP:conf/hotpar/RubinsteynHWS12}:

@@

We should explain the classification of these runtimes

@@

Except GraalPython, one can notice 3 main clusters.
The first cluster contains the reference implementations (CPython2 and CPython3) and the implementations that are based on interpreters (IronPython, IntelPython, ActivePython, Jython, and Nuitka).
Moreover, each implementation is the closest to its reference Python version.
The interpreters that are based on other virtual machines, such as Jython with JVM and IronPython with .NET, behave slightly differently from the reference implementation.
As for the MicroPython, It is the furthest from the references due to his behavior toward exceptions.
The second cluster groups the interpreter implementations that are based on a JIT compiler (PyPy2 and PyPy3).
Closer to them we find Shedskin, which is a translator that converts the Python code into C++ code to be later compiled into binary code.
Table~\ref{table:manwithyu} shows that this cluster is by far the most energy-efficient one.
While Numba2 and Numba3 are also JIT libraries, they differ from PyPy due to their manual optimization.
Unlike PyPy, which will decide by itself which part of Python to optimize.
In Numba, the developer must target the portion of the code that will be passed to JIT, while the rest of the code will be executed by the regular interpreter (aka CPython).

\paragraph{Remark:}
Because GraalPython was still in its early stages, some techniques required an abnormally lengthy time, affecting the clustering algorithm.

@@

\begin{figure}[!htb]
      \centering
      \includegraphics[width=1.0\linewidth]{imgs/alltomti_performance}
      \caption{Different interpreter's energy scores.}
      \label{fig:tommi_all}
\end{figure}

@@

\subsection{Conclusion}
One can conclude that the choice of Python interpreter has a significant impact on the programs' energy consumption.
This investigation is made more intriguing by the absence of a universal solution.
The primary downside is the incompatibility of some of these solutions, which causes us to make concessions when we need a generic answer.

>>>>

      \item \textbf{Parakeet}~\cite{rubinsteyn2012parakeet}:

@@

% We should explain the classification of these runtimes

@@

Except GraalPython, one can notice 3 main clusters.
The first cluster contains the reference implementations (CPython2 and CPython3) and the implementations that are based on interpreters (IronPython, IntelPython, ActivePython, Jython, and Nuitka).
Moreover, each implementation is the closest to its reference Python version.
The interpreters that are based on other virtual machines, such as Jython with JVM and IronPython with .NET, behave slightly differently from the reference implementation.
As for the MicroPython, It is the furthest from the references due to his behavior toward exceptions.
The second cluster groups the interpreter implementations that are based on a JIT compiler (PyPy2 and PyPy3).
Closer to them we find Shedskin, which is a translator that converts the Python code into C++ code to be later compiled into binary code.
Table~\ref{table:manwithyu} shows that this cluster is by far the most energy-efficient one.
While Numba2 and Numba3 are also JIT libraries, they differ from PyPy due to their manual optimization.
Unlike PyPy, which will decide by itself which part of Python to optimize.
In Numba, the developer must target the portion of the code that will be passed to JIT, while the rest of the code will be executed by the regular interpreter (aka CPython).
\paragraph{Remark:}
Because GraalPython was still in its early stages, some techniques required an abnormally lengthy time, affecting the clustering algorithm.

@@

\begin{figure}[!htb]
      \centering
      \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{imgs/alltomti_performance}
      \caption{Different interpreter's energy scores.}
      \label{fig:tommi_all}
\end{figure}

@@

\subsection{Conclusion}
One can conclude that the choice of Python interpreter has a significant impact on the programs' energy consumption.
While most of the candidates used Python2, we observed that different strategies that were implemented within each candidate have a significant impact on the energy consumption. 
The major improvement of the energy was related to the use of JIT, which is a technique that compiles the code into machine code at runtime. 
The effect of the compilation of the python code into machine code was also confirmed with the use of translators such as shedskin.

This investigation is made more intriguing by the absence of a universal solution.
The primary downside is the incompatibility of some of these solutions, which causes us to make concessions when we need a generic answer.
